%% ============================================================
%% main.tex — PNAS research article format
%% Requires pnas-new.cls and pnas-new.bst (both in paper/)
%% Compile: pdflatex main && bibtex main && pdflatex main && pdflatex main
%% ============================================================

\documentclass[9pt,twoside,lineno]{pnas-new}

\templatetype{pnasresearcharticle}

%% ---- Title --------------------------------------------------
\title{Overconfidence in Confidence Interval Estimates Produced by
       Frontier Large Language Models}

%% ---- Authors ------------------------------------------------
\author[a,1]{Author Placeholder}

\affil[a]{Institution Placeholder}

\leadauthor{Walters}

%% ---- Significance statement ---------------------------------
\significancestatement{%
Large language models (LLMs) are increasingly deployed as
decision-support tools, yet their capacity to express calibrated
uncertainty has not been systematically evaluated. We queried three
frontier LLMs across three studies covering equity markets (1 to 22-day
horizons) and six heterogeneous domains. All models were substantially
overconfident, with GPT-4 the most miscalibrated and Claude the least
across all studies. Critically, miscalibration is not uniform: all
models overstate certainty in commodity and equity markets while
overstating uncertainty in cryptocurrency and forex---a domain-confidence
inversion consistent with domain-specific priors inherited from training
corpora. These findings show that global calibration benchmarks are
insufficient; domain-stratified evaluation is necessary to characterize
model-specific calibration risk in high-stakes applications.%
}

%% ---- Author declarations ------------------------------------
\authordeclaration{The authors declare no competing interests.}
\equalauthors{\textsuperscript{1}Correspondence:
  author\texttt{@}institution.edu}
\keywords{large language models | calibration | overconfidence |
  interval estimation | forecasting}

%% ---- Abstract -----------------------------------------------
\begin{abstract}
We document systematic miscalibration in confidence interval estimates
produced by three frontier large language models (Claude, GPT-4, Gemini).
In Study~1, all three models were substantially overconfident when
predicting 1-day equity prices, with GPT-4 exhibiting the largest
Expected Calibration Error (ECE\,=\,43.7\%) and Claude the smallest
(ECE\,=\,19.7\%). In Study~2, using 100 equities across nine prediction
horizons spanning 1--22 days ($N > 18{,}000$ scored predictions), Claude
and GPT-4 became increasingly miscalibrated at longer horizons, whereas
Gemini maintained near-target coverage at the longest horizon tested
(ECE\,=\,2.5\% at 22~days). In Study~3, extending to six heterogeneous
prediction domains, we found a systematic domain-confidence inversion:
all models were overconfident in commodity markets
($\mu = 0.25$ uniformly) but underconfident in cryptocurrency
($\mu = 1.35$--$14.4$), while achieving near-target calibration on
weather. Across all three studies, GPT-4 was the most overconfident
model and Claude the most calibrated. These findings show that LLM
miscalibration is domain- and horizon-structured rather than uniform,
with direct implications for deploying LLMs in high-stakes decision
support.
\end{abstract}

\additionalelement{}
\dates{This manuscript was compiled on \today}
\doi{10.1073/pnas.XXXXXXXXXX}
\acknow{The authors thank the maintainers of the open data APIs used in this
study: yfinance, CoinGecko, Open-Meteo, and ESPN. No external funding
was received for this research.}

%% ============================================================
\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}
  {\ifthenelse{\boolean{singlecolumn}}
    {\abscontentformatted}{\abscontent}}{}

%% ---- Body sections -----------------------------------------
\input{sections/introduction}
\input{sections/study1}
\input{sections/study2}
\input{sections/study3}
\input{sections/discussion}

%% ---- Acknowledgements + Bibliography -----------------------
\showacknow
\bibliography{references}

\end{document}
